---
title: "crudr: Introduction"
author: "Janko Thyson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{README}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(crudr)
run <- FALSE
root_dir <- paste(rep("../", 1), collapse = "")
knitr::opts_chunk$set(comment = "#>", collapse = TRUE)
knitr::opts_knit$set(root.dir = root_dir)
```

# UML

```{r echo=FALSE}
crudr_uml <- '
  @startuml
  ICrud *-- Crud: implements <
  Crud <|-- Crud.Settings: extends <

  class ICrud {
    +init()
    +has()
    +create()
    +read()
    +update()
    +delete()
    +reset()
    -stopIfInterface()
  }
  class Crud {
    +main: environment
    -initial_state: list
    -stopIfEmpty()
    -createMessage(),
    -cacheInitialState()
  }
  class Crud.Settings {
    +main: function
    -cacheInitialState()
  }
  @enduml
'
# knitr::opts_knit$set(root.dir = root_dir)
uml_file <- crudr:::plantuml3(
  crudr_uml, 'Architecture', 
  filename = "vignettes/uml",
  jarfile = "lib/plantuml.jar",
  normalize = TRUE
)
# print(uml_file)
# uml_file <- normalizePath(uml_file)
# print(uml_file)
# print(getwd())
# knitr::opts_knit$set(root.dir = getwd())
```

![UML](`r uml_file`)

# The interface

Class `ICrud` serves as the inferface class that can be **implemented** by other classes. It defines nothing but the publicly available methods. 

As no explicit instances of this class should be created/used, calling the its methods results in an error informing the user that he is working with the interface itself instead of an implementation of the interface.

```{r, eval = run}
inst <- ICrud$new()
try(inst$init())
try(inst$has())
try(inst$create())
try(inst$read())
try(inst$update())
try(inst$delete())
try(inst$reset())
```

# The base class

Class `Crud` implements the interface `ICrud` and serves as a base clase for inheritance.

```{r, eval = run}
inst <- Crud$new()
class(inst)
inherits(inst, "ICrud")
```

## Getters/setters

```{r, eval = run}
inst <- Crud$new(main = as.environment(list(a = 1, b = 2)))

value <- inst$getMain()
as.list(value, sorted = TRUE)

inst$setMain(as.environment(list(a = 10, b = 20)))
inst$getMain()$a
inst$getMain()$b
```

## CRUD methods

### Instantiate 

```{r, eval = run}
inst <- Crud$new()
```

### Init

Provide initial values.

Note the flexible format by which you can provide input.

```{r, eval = run}
inst$init(a = 1, b = 2, "c/foo/bar" = 3, list(d = 4, "e/foo/bar" = 5))
```

### Has 

Check if values exist based on one or multiple keys:

```{r, eval = run}
inst$has("a")
inst$has("c/foo")
inst$has("a", "b")
inst$has("a", "e/foo/bar")

inst$has("c")
inst$has("a", "c", "c/bar")
```

### Create

Create one or multiple key-value pairs. Takes into consideration existing key-value pairs which, by default, are not overwritten. Overwrites need to be explicitly authorized by `overwrite = TRUE`. 

Note the different levels of strictness that can be applied. This is something that most methods support.

```{r, eval = run}
inst$create(f = TRUE)
inst$create(g = TRUE, h = letters[1:3])
inst$create("i/foo" = 1, "c/foo/bar/new" = 1)
## TODO 2015-11-04: review behavior for preserving values; turning everything
## into a list would probably be nicer as would happen when calling
## `inst$create("i/foo" = 1, "c/foo/bar/new/something" = 1)`
as.list(inst$getMain(), sorted = TRUE)

(inst$create(c = 30, "c/foo" = 30))
## --> c already existed with value = 3
as.list(inst$getMain(), sorted = TRUE)
inst$create(c = 30, strict = 1)
try(inst$create(c = 30, strict = 2))
try(inst$create(c = 30, strict = 3))

## Overwrite //
inst$create(c = 30, "c/foo" = 30, overwrite = TRUE)
## TODO 2015-11-04: review behavior for preserving values; turning everything
## into a list would probably be nicer
as.list(inst$getMain(), sorted = TRUE)
## --> c overwritten
```

### Read

Read/get values by one or multiple keys. Non-existing key-value pairs will result in return values of NULL.

Strictness levels `0` through `3` apply.

```{r, eval = run}
inst$read()
inst$read("a")
inst$read("a", "b", "c/foo")

inst$read("a", "b", "x")
try(inst$read("a", "b", "x", strict = 3))
```

### Update

Update values by one or multiple keys. 

Strictness levels `0` through `3` apply.

```{r, eval = run}
inst$read("a")
inst$update(a = 10)
inst$read("a")

inst$read("a", "b")
inst$update(a = 1000, b = 2000)
inst$read("a", "b")

inst$update(x = 99)
inst$read("x")

inst$read("a", "x")
inst$update(a = 1, x = 99)
try(inst$update(a = 1, x = 99, strict = 3))
inst$read("a", "x")
```

### Delete

Delete key-value pairs based on keys. Only existing pairs can be deleted. 

Strictness levels `0` through `3` apply.

```{r, eval = run}
inst$read()
inst$delete("f")

inst$delete("d", "c")
inst$read()

inst$delete("x")
inst$read()
inst$delete("b", "x")
try(inst$delete("b", "x", strict = 3))
inst$read()
```

### Reset

```{r, eval = run}
inst$reset()
inst$read()

inst$reset(type = "hard")
inst$read()
```
